---
title: "CSB Project Code Description"
author: "Cody Cole"
date: "3/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Shell Script

This script is found in the file `sandbox/file_movement_script.sh`. The purpose of these commands are to copy the data that are saved on a flashdrive, which is mimicked here by the folder `flashdrive` that has the files stored under `flashdrive/Growth_Curves` and create a csv for the data in the `input` folder that will be used to add the sample names and conditions to the data. When running this script, it will move the data files, but it won't change the csv files for labels as I have already run that code and filled out the labels for the purpose of running the R script. 

I attempted to make the regular expressions pretty general to account for the differences that may arise from my file naming, and it is written in a way to prevent overrighting the files that may already be present. 

```{bash}
#!/bin/bash

#Change permissions to copy and rename files
chmod a+rwx ../flashdrive/Growth_Curves/*

#Searches the flashdrive for files that contain the word test
test_files=$(find ../flashdrive/Growth_Curves/ -iname '*test*')

#For loop that takes each file from the flashdrive and changes the name copies it to the data folder 
for file in $test_files 
do 
  #The sed function here takes all the information past test (est to be case insensitve) to rename the file
  file_tag=$(echo "$file" | sed -ne 's/.*est//p')
  #In this case it does the same, but leaves out the .xlsx from the file name
  file_tag_without_xlsx=$(echo "$file" | sed -ne 's/.*est//;s/\..*$//p')
  #Copies the file to data folder without overrighting files already copied
  cp -n $file ../data/Nanobody_Slay_Test"$file_tag"
  #Creates a label csv that will be used to match to the data file in the R script based on a template in folder
  cp -n ../input/template.csv ../input/Nanobody_Slay_Test"$file_tag_without_xlsx"_Labels.csv
done


echo "Now go to the input file directory and fill out the labels for your tests"
```


## R Script

This R script from `sandbox/code_script.R` is responsible for generating descriptive plots for all the files that are inserted into the `data` folder.

```{r}
library(tidyverse)
library(readxl)
library(stringi)
```

Here the files in the `data` folder are amended to `data_files` which will be used for the **for** loop. `plot_directory` contains the what has already been analyzed. 

```{r}
#Access files located in the data and plot directories 
data_files <- dir("../data", 
                  recursive=TRUE,full.names=TRUE)

plot_directory <- dir("../plots/",
                      recursive = FALSE,
                      full.names = FALSE)
```

This **for** loop goes through each file in the `data` to tidy the data and generate two plots. There is the `overview_plot.pdf` looks at the entire 96-well plate to get a general idea of what went on with the assay, and the `main_plot.pdf` is the graph that compiles the replicates and displays the data in a much more readable way. 

There are a few **if** statments imbeded to prevent reanalyzing data that has already been process, progress the for loop in the event it comes across an empty label file, set desired y-limits, and change the `main_plot.pdf` output for optimal display based on the data. 

The output plots are stored under their test number in the `plot` folder. 

```{r}
#For loop that cleans and generates graphs for 
for (file in data_files) {
  
  #If statement to prevent reanalyzing already processed data
  #To rerun already analyzed data, delete files located in plots folder
  if (sum(str_detect(file, plot_directory)) == 0) {
    
    #Load in the data from files specified in for loop
    growth_curve_data <- read_excel(file)
    
    #Extracts test number from file to match to the label file that will be joined
    test_number <- stri_sub(file, 23, -6)
    growth_curve_labels <- read_csv(paste0("../input/Nanobody_Slay_", test_number,"_Labels.csv"))
    
    #This is an if statement to progress for loop in the event labels were not added to the label file
    if (length(unique(growth_curve_labels$Plasmid)) <= 1) next()
    
    
    #Clean up the data and join the labels that have to be input by the user
    data_columns <- ncol(growth_curve_data)
    growth_curve_data_final <- growth_curve_data %>%
      rename(Temperature = `Temperature(¡C)`) %>% 
      gather(Well, Turbidity, 3:data_columns) %>% 
      separate(Time, c("Date", "Time"), sep = " ") %>%
      select(-Date) %>% 
      separate(Time, c("Hours", "Minutes"), extra = "drop", remove = FALSE) %>% 
      type_convert(cols(Hours = col_integer(),
                        Minutes = col_integer())) %>% 
      mutate(Hours = (Hours + (Minutes / 60))) %>% 
      select(-Minutes) %>% 
      left_join(growth_curve_labels) %>% 
      #This unite function creates unique identifer for the samples that includes their replicates for overview_plot
      unite(Sample_id, 
            Strain, Plasmid, Media, Condition1, Condition2, 
            sep = "  ", 
            remove = FALSE,
            na.rm = TRUE) %>% 
      unite(Condition1,
            Media, Condition1, 
            sep = " + ", 
            remove = FALSE,
            na.rm = TRUE) %>% 
      unite(Strain_Plasmid,
            Strain, Plasmid, 
            sep = " + ", 
            remove = FALSE,
            na.rm = TRUE)
      
    
    #Summary of data for building the main_plot (This throws warning cause of NAs in Blank)
    growth_curve_summary <- growth_curve_data_final %>%
      group_by(Time, Hours, Temperature, Strain, Plasmid, Media, Condition1, Condition2, Strain_Plasmid) %>% 
      summarise(Mean_Turbidity = mean(Turbidity),
                SD = sd(Turbidity), 
                N = n(), 
                SEM = SD/sqrt(N),
                CI = SEM * qt(0.975, N-1)) %>% 
      group_by(Hours, Media, Condition1, Condition2) %>%
      mutate(Mean_Turbidity = Mean_Turbidity - Mean_Turbidity[Strain == "Blank"]) %>% 
      filter(Strain != "Blank") %>% 
      mutate(y_upper_limit = (Mean_Turbidity + CI)) %>% 
      mutate(y_lower_limit = (Mean_Turbidity - CI))
    
    
    #Set the limits used for ploting
    x_upper <- max(growth_curve_summary$Hours) + .1
    y_upper <- max(growth_curve_summary$y_upper_limit)
    y_lower <- min(growth_curve_summary$y_lower_limit)
    
    
    #Set preferred y limits for easy comparison of graphs between different tests
    if (y_upper < 1.75) {
      y_upper <- 1.75
    } 
    if (y_lower >= 0) {
      y_lower <- 0
    } 
    
    
    #This factors variables that are used for plotting and adjusts the order based 
    #the input of the user by using the unique function in levels
    growth_curve_summary$Strain_Plasmid <- factor(growth_curve_summary$Strain_Plasmid, 
                                           levels = c(unique(growth_curve_data_final$Strain_Plasmid)))
    growth_curve_summary$Condition1 <- factor(growth_curve_summary$Condition1, 
                                              levels = c(unique(growth_curve_data_final$Condition1)))
    growth_curve_summary$Condition2 <- factor(growth_curve_summary$Condition2, 
                                              levels = c(unique(growth_curve_data_final$Condition2)))
    growth_curve_data_final$Well <- factor(growth_curve_data_final$Well, 
                                           levels = c(unique(growth_curve_data_final$Well)))
    
    
    #Create the directory that the plots will go into dempending on the test number of the file
    dir.create(paste0("../plots/", test_number))
    
    
    #Generate overview plot for the whole 96-well plate
    ggplot(data = growth_curve_data_final) + 
      aes(x = Hours, y = Turbidity, color = Sample_id) +
      geom_line(size = 1) +
      facet_wrap(.~Well, ncol = 12) + 
      ylab(expression("Turbidity (OD"[600]*")")) +
      theme(legend.position = "none") +
      scale_x_continuous(breaks = seq(0, x_upper, 5)) +
      scale_y_continuous(breaks = seq(0, y_upper, .75), 
                         limits = c(y_lower, y_upper)) +
      ggsave(paste0("../plots/", test_number,"/overview_plot.pdf"), 
             device = "pdf",
             width = 30,
             height = 18,
             units = "cm")
    
    
    #This if statement changes the main plot that is generated based on whether or not Media is different between the samples
    #In the event it is, it looks better when plotted differently 
    if (length(unique(growth_curve_summary$Media)) > 1) {
      
      #generates main_plot in  special case  
      ggplot(data = growth_curve_summary) + 
        aes(x = Hours, y = Mean_Turbidity, color = Strain_Plasmid) +
        geom_point() +
        geom_line() +
        geom_errorbar(aes(ymin = (Mean_Turbidity - CI), 
                          ymax = (Mean_Turbidity + CI), width = .18)) + 
        facet_grid(Condition2~Condition1) + 
        ylab(expression("Turbidity (OD"[600]*")")) + 
        theme_bw() +
        theme(legend.position = "bottom",
              legend.title = element_blank()) +
        scale_y_continuous(expand = c(.02, 0), 
                           breaks = seq(0, y_upper, .25), 
                           limits = c(y_lower, y_upper)) + 
        scale_x_continuous(expand = c(.015, 0), 
                           breaks = seq(0, x_upper, 1), 
                           limits = c(0, x_upper)) +
        guides(color = guide_legend(nrow=3,byrow=TRUE)) +
        ggsave(paste0("../plots/", test_number,"/main_plot.pdf"), 
               device = "pdf",
               width = 30,
               height = 18,
               units = "cm")
      
    } else {
      
      #generates main_plot for rest of the data
      ggplot(data = growth_curve_summary) + 
        aes(x = Hours, y = Mean_Turbidity, color = Strain_Plasmid) +
        geom_point() +
        geom_line() +
        geom_errorbar(aes(ymin = (Mean_Turbidity - CI), 
                          ymax = (Mean_Turbidity + CI), width = .18)) + 
        facet_grid(Condition1~Condition2) + 
        ylab(expression("Turbidity (OD"[600]*")")) + 
        theme_bw() +
        theme(legend.position = "bottom",
              legend.title = element_blank()) +
        scale_y_continuous(expand = c(.02, 0), 
                           breaks = seq(0, y_upper, .25), 
                           limits = c(y_lower, y_upper)) + 
        scale_x_continuous(expand = c(.015, 0), 
                           breaks = seq(0, x_upper, 1), 
                           limits = c(0, x_upper)) +
        guides(color = guide_legend(nrow=3,byrow=TRUE)) +
        ggsave(paste0("../plots/", test_number,"/main_plot.pdf"), 
               device = "pdf",
               width = 30,
               height = 18,
               units = "cm")
    } 
    
  }
}
```


